
```{r}
## Create figures for FIG 1/2 -- EXCITATORY NEURONS
```


```{r cars}
library(Seurat) 
#library(SeuratDisk)
#library(SeuratData)
library(ggplot2)
library(patchwork)
library(dplyr)
#library(sceasy)
library(reticulate)
#library(rliger)
library(magrittr) # needs to be run every time you start R and want to use %>%
library(dplyr) 
library(Seurat)
library(tidyverse)
library(SeuratWrappers)
library(harmony)
library(batchelor)
library(usethis)
library(Seurat)
library(monocle3)
library(dplyr)
```


```{r}
all_data <- readRDS("all_data.rds")
```




```{r}
all_data = FindNeighbors(all_data)
all_data = FindClusters(all_data, resolution = 2)
# Set 'RNA_snn_res.0.3' as the active identity class
```


```{r}
DefaultAssay(all_data) <- "RNA"
Idents(all_data) <- "RNA_snn_res.2"
VlnPlot(
  all_data,
  features = "En1",
  group.by = "seurat_clusters"
)
VlnPlot(
  all_data,
  features = "En2",
  group.by = "seurat_clusters"
)
```
```{r}
URL <- subset(
  all_data,
  subset = seurat_clusters %in% c(
    "26", "27", "24", "3", "28", "36",
    "38", "10", "21", "31", "39", "30",
    "29", "35", "5"
  )
)
```





```{r}
DimPlot(URL,label = TRUE)
```




```{r}

URL.mnn <- RunFastMNN(object.list = SplitObject(URL, split.by = "Sample"), assay = "RNA", features = 5000, verbose = FALSE)
URL[["mnn"]] <- CreateDimReducObject(Embeddings(URL.mnn, reduction = "mnn"), assay = "RNA")

```

```{r}
URL <- FindNeighbors(URL, reduction = "mnn", dims = 1:50)

URL <- RunUMAP(URL, reduction = "mnn", dims = 1:50, spread = 3, min.dist = 0.3)
```


```{r}
DimPlot(URL, reduction = "umap", group.by = 'TimePoint') + NoAxes()
DimPlot(URL, reduction = "umap", group.by = 'branch', label = TRUE) + NoAxes()
FeaturePlot(URL, feature = "Pax6", max.cutoff = 2)
FeaturePlot(URL, feature = "Lhx9", max.cutoff = 2)

```

```{r}
p <- DimPlot(URL, group.by = "Sample")
p

ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/URL_Sample.png",
  plot = p,
  width = 7,
  height = 5,
)
```


```{r}
URL <- FindNeighbors(URL, reduction = "mnn")
URL <- FindClusters(URL, resolution = 2)
DimPlot(URL, group.by = "RNA_snn_res.2", label = TRUE)


```

```{r}
DefaultAssay(URL) <- "RNA"
URL <- FindClusters(URL, resolution = 1.5)

```


```{r}
URL$collector <- "Butts"

p <- DimPlot(URL, group.by = "collector")
p

ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/URL_collector.png",
  plot = p,
  width = 7,
  height = 5,
)
```

```{r}
FeaturePlot(URL, features = "Reln", max.cutoff = 2)
FeaturePlot(URL, features = "Zfhx3", max.cutoff = 2)
FeaturePlot(URL, features = "Meis2", max.cutoff = 2)
FeaturePlot(URL, features = "Lmx1a", max.cutoff = 2)

```

```{r}
FeaturePlot(URL, features = "Nr2f1", max.cutoff = 2)
FeaturePlot(URL, features = "Isl1", max.cutoff = 2)
FeaturePlot(URL, features = "Casz1", max.cutoff = 2)

FeaturePlot(URL, features = "Pcp4", max.cutoff = 2)
FeaturePlot(URL, features = "Nfia", max.cutoff = 2)

```


```{r}
URL$cell_type_coarse <- recode(URL$RNA_snn_res.1.5,
    "0" = "Interposed/LateralDCN", 
    "1" = "Granule Cells+UBCs",
    "2" = "Extracerebellar-fated", 
    "3" = "RL",
    "4" = "Granule Cells+UBCs", 
    "5" = "RL",
    "6" = "Granule Cells+UBCs", 
    "7" = "Extracerebellar-fated",
    "8" = "Extracerebellar-fated", 
    "9" = "Extracerebellar-fated",
    "10" = "Granule Cells+UBCs", 
    "11" = "Granule Cells+UBCs",
    "12" = "Extracerebellar-fated",
    "13" = "Extracerebellar-fated", 
    "14" = "RL",
    "15" = "Granule Cells+UBCs", 
    "16" = "Interposed/LateralDCN",
    "17" = "RL", 
    "18" = "Medial DCN",
    "19" = "Extracerebellar-fated",
    "20" = "Interposed/LateralDCN", 
    "21" = "Medial DCN",
    "22" = "Interposed/LateralDCN", 
    "23" = "Granule Cells+UBCs",
    "24" = "Extracerebellar-fated", 
    "25" = "Granule Cells+UBCs",
    "26" = "Extracerebellar-fated", 
    "27" = "Extracerebellar-fated"
)

```


```{r}
DimPlot(URL, group.by = "cell_type_coarse") + NoAxes()
```
```{r}
colors = c("Medial DCN"="blue", "Granule Cells+UBCs"="purple", "Interposed/LateralDCN"="orange", "Extracerebellar-fated"="black", "RL" = "lightblue")

p <-DimPlot(URL, group.by = "cell_type_coarse", cols = colors) + NoAxes()

ggsave("C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/cerebellum_2branch_branches_cncombined.svg", plot = p, width = 8, height = 6, units = "in", dpi = 300)

colors = c("Medial DCN"="blue", "Granule Cells+UBCs"="purple", "Interposed/LateralDCN"="blue", "Extracerebellar-fated"="black", "RL" = "lightblue")

p <-DimPlot(URL, group.by = "cell_type_coarse", cols = colors) + NoAxes()

ggsave("C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/cerebellum_2branch_branches_cncombined.png", plot = p, width = 8, height = 6, units = "in", dpi = 300)


p <-DimPlot(URL, group.by = "cell_type_coarse", cols = colors) + NoAxes()

ggsave("C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/cerebellum_2branch_branche_cncombineds.png", plot = p, width = 8, height = 6, units = "in", dpi = 300)
p
```

```{r}
all_markers <- FindAllMarkers(
  URL,
  group.by = "cell_type_coarse",
  only.pos = TRUE,
  min.pct = 0.75,
  logfc.threshold = 0.25
)
```


```{r}
# 5. Get top 10 genes per cluster
top10_markers <- all_markers %>%
  group_by(cluster) %>%
  slice_max(n = 5, order_by = avg_log2FC)

# 6. View results
head(top10_markers)
```


```{r}
top10_markers_unique <- unique(as.character(top10_markers$gene))
p <- DotPlot(
  URL,
  features = top10_markers_unique,
  group.by = "cell_type_coarse",
  assay = "RNA",
  cols = c("yellow", "purple"),
  dot.scale = 10,
  scale.by = "radius",
  col.min = 0,
  col.max = 1
) +
  scale_fill_gradient2(
    low = "yellow",
    mid = "yellow",
    high = "purple",
    midpoint = 0
  ) +
  RotatedAxis() +
  theme(
    axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5,
      family = "Arial"
    ),
    axis.text.y = element_text(family = "Arial"),
    text = element_text(family = "Arial")
  )

p
```


```{r}
ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/mouse_exc_dotplot.svg",
  plot = p,
  width = 12,
  height = 5,
)
```



```{r}
URL <- FindClusters(URL, resolution = 1)
```


```{r}
URL$branch <- recode(URL$RNA_snn_res.1,
    "0" = "early", 
    "1" = "transition",
    "2" = "early", 
    "3" = "transition",
    "4" = "early", 
    "5" = "late",
    "6" = "transition", 
    "7" = "early",
    "8" = "early", 
    "9" = "early",
    "10" = "early", 
    "11" = "late",
    "12" = "late",
    "13" = "late", 
    "14" = "early",
    "15" = "late", 
    "16" = "late",
    "17" = "early", 
    "18" = "late",
    "19" = "late",
    "20" = "early"
  
    # Add more as needed
)
```

```{r}
DimPlot(URL, group.by = "branch")

```
```{r}
URL <- subset(URL, Mki67 == 0)
URL <- subset(URL, Top2a == 0)

```

```{r}
saveRDS(URL, "URL_082026.rds")
```

```{r}
URL <- readRDS("URL_082026.rds")
```


```{r}
p <- DimPlot(URL, reduction = "umap", group.by = 'TimePoint') + NoAxes()
ggsave("C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/allURL_timepoint.png", plot = p, width = 8, height = 6, units = "in", dpi = 300)

p <- DimPlot(URL, reduction = "umap", group.by = 'Sample') + NoAxes()
ggsave("C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/allURL_sample.png", plot = p, width = 8, height = 6, units = "in", dpi = 300)

p <- FeaturePlot(URL, reduction = "umap", features = 'Pax6') + NoAxes()
ggsave("C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/allURL_Pax6.png", plot = p, width = 8, height = 6, units = "in", dpi = 300)

p <- FeaturePlot(URL, reduction = "umap", features = 'Lhx9') + NoAxes()
ggsave("C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/allURL_Lhx9.png", plot = p, width = 8, height = 6, units = "in", dpi = 300)
```


```{r}
FeaturePlot(URL, features = "Pax6")
FeaturePlot(URL, features = "Lhx9")

DimPlot(URL, group.by = "branch")
```


```{r}
# Initialize new metadata column in URL
URL$branch <- as.character(URL$branch)

URL$med_2p_branch_full <- URL$branch

URL$med_2p_branch_full <- recode(
  URL$branch,
  "early"  = "Midbrain-fated Cells", 
  "transition" = "RL",
  "late" = "Granule Cells+UBCs"
)
# Remove "_n" only when it appears at the end of the name
colnames(combined_alltp) <- sub("_1", "", colnames(combined_alltp))

# Now recompute the shared cells, if appropriate:
cells_shared <- intersect(colnames(URL), colnames(combined_alltp))

cells_shared <- intersect(colnames(URL), colnames(combined_alltp))

URL$med_2p_branch_full[cells_shared] <-
  combined_alltp$med_2p_branch[cells_shared]


```


```{r}
# Initialize new metadata column in URL
URL$branch <- as.character(URL$branch)

URL$coarse_populations <- URL$branch

URL$coarse_populations <- recode(
  URL$branch,
  "early"  = "Midbrain-fated Cells", 
  "transition" = "RL",
  "late" = "Granule Cells+UBCs"
)
# Remove "_n" only when it appears at the end of the name
colnames(combined_alltp) <- sub("_1", "", colnames(combined_alltp))

# Now recompute the shared cells, if appropriate:
cells_shared <- intersect(colnames(URL), colnames(combined_alltp))

cells_shared <- intersect(colnames(URL), colnames(combined_alltp))

URL$coarse_populations[cells_shared] <-
  "Cerebellar Nuclei"


```



```{r}
library(Seurat)
library(dplyr)

all_markers <- FindAllMarkers(
  URL,
  group.by = "branch",
  only.pos = TRUE,        # keep only upregulated genes
  min.pct = 0.25,         # expressed in ≥25% cells
  logfc.threshold = 0.25  # minimum log fold change
)

head(all_markers)

all_markers <- all_markers %>%
  filter(p_val_adj < 0.05)

top10 <- all_markers %>%
  group_by(cluster) %>%
  slice_max(n = 5, order_by = avg_log2FC)

top10_ordered <- top10 %>%
  arrange(cluster, desc(avg_log2FC))

top_genes <- top10_ordered %>%
  pull(gene) %>%
  unique()

URL$branch <- factor(URL$branch,levels = rev(c("late",
                                "transition",
                                "early"
                                
                              )))


p <- DotPlot(
  URL,
  features = top_genes,
  assay = "RNA",
  group.by = 'branch',
  cols = c('yellow', 'purple'),
  dot.min = -0.6,
  dot.scale = 10,
  col.min = -1,
  col.max = 2
  #min.pct = 0.25
) + 
  scale_fill_gradient2(
    low = "yellow",
    mid = "yellow",
    high = "purple",
    midpoint = 0
  ) +
  RotatedAxis() +
  theme(
    axis.text.x = element_text(
      angle = 90,   # vertical text
      hjust = 1,
      vjust = 0.5,
      family = "Arial"
    ),
    axis.text.y = element_text(family = "Arial"),
    text = element_text(family = "Arial")
  )

ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/allURL_dotplot_branch.svg",
  plot = p,
  width = 9,
  height = 5,
  device = svglite::svglite
)
```
```{r}
p
```



```{r}

branch_pal <- c(
  "Int/Lat Prog" = "#fca800",
  "Med"          = "blue",
  "IntA/Lat"     = "#fca800",
  "IntA"     = "#fca800",
  "Lat"     = "#fca800",
  "RL"           = "lightblue",
  #"IntP/IntA"  
  "IntP"         = "#fca800",
  #"Med_gly"      = "#17becf",
  "Early Medial"        = "blue",
  "Late Medial"         = "blue",
  "Midbrain-fated Cells"  = "black",
  "Granule Cells+UBCs" = "purple"
)

# Make sure the column is a factor
URL$med_2p_branch_full <- factor(URL$med_2p_branch_full)
# cells for the clusters you want in the back
back_clusters <- c("Midbrain-fated Cells", "Granule Cells+UBCs")

# order cells: back clusters first, then all others
cell_order <- c(
  rownames(URL@meta.data[URL$med_2p_branch_full %in% back_clusters, ]),
  rownames(URL@meta.data[!(URL$med_2p_branch_full %in% back_clusters), ])
)

# store as new identity factor (optional)
URL$med_2p_branch_full_ordered <- factor(
  URL$med_2p_branch_full[cell_order],
  levels = unique(URL$med_2p_branch_full[cell_order])
)
```

```{r}
URL$cell_type_coarse <- recode(URL$RNA_snn_res.1.5,
    "0" = "Interposed/Lateral DCN", 
    "1" = "Granule Cells+UBCs",
    "2" = "Extracerebellar-fated", 
    "3" = "RL",
    "4" = "Granule Cells+UBCs", 
    "5" = "RL",
    "6" = "Granule Cells+UBCs", 
    "7" = "Extracerebellar-fated",
    "8" = "Extracerebellar-fated", 
    "9" = "Extracerebellar-fated",
    "10" = "Granule Cells+UBCs", 
    "11" = "Granule Cells+UBCs",
    "12" = "Extracerebellar-fated",
    "13" = "Extracerebellar-fated", 
    "14" = "RL",
    "15" = "Granule Cells+UBCs", 
    "16" = "Interposed/Lateral DCN",
    "17" = "RL", 
    "18" = "Medial DCN",
    "19" = "Extracerebellar-fated",
    "20" = "Interposed/Lateral DCN", 
    "21" = "Medial DCN",
    "22" = "Interposed/Lateral DCN", 
    "23" = "Granule Cells+UBCs",
    "24" = "Extracerebellar-fated", 
    "25" = "Granule Cells+UBCs",
    "26" = "Extracerebellar-fated", 
    "27" = "Extracerebellar-fated"
  
  
    # Add more as needed
)
colors = c("Medial DCN"="blue", "Granule Cells+UBCs"="purple", "Interposed/Lateral DCN"="orange", "Extracerebellar-fated"="black", "RL" = "lightblue")

p <-DimPlot(URL, group.by = "cell_type_coarse",cols = colors, label = FALSE) + NoAxes()

ggsave(
  "excitatory_mouse_umaps/URLcell_type.png",
  plot = p,
  width = 7,
  height = 5,
)
```


```{r}
p
```


```{r}
cell_order <- rev(c(
  "Granule Cells+UBCs",
  "Medial DCN",
  "Interposed/Lateral DCN",
  "Extracerebellar-fated",
  "RL"
))

URL$cell_type_coarse <- factor(
  URL$cell_type_coarse,
  levels = cell_order
)

all_markers <- FindAllMarkers(
  URL,
  group.by = "cell_type_coarse",
  only.pos = TRUE,
  min.pct = 0.5,
  logfc.threshold = 0.5
)

# 5. Get top 10 genes per cluster
top10_markers <- all_markers %>%
  group_by(cluster) %>%
  slice_max(n = 5, order_by = avg_log2FC)

# 6. View results
head(top10_markers)

p <- DotPlot(
  URL,
  features = top10_markers$gene,
  assay = "RNA",
  group.by = 'cell_type_coarse',
  cols = c('yellow', 'purple'),
  dot.min = -0.6,
  dot.scale = 10,
  col.min = -1,
  col.max = 2
  #min.pct = 0.25
) + 
  scale_fill_gradient2(
    low = "yellow",
    mid = "yellow",
    high = "purple",
    midpoint = 0
  ) +
  RotatedAxis() +
  theme(
    axis.text.x = element_text(
      angle = 90,   # vertical text
      hjust = 1,
      vjust = 0.5,
      family = "Arial"
    ),
    axis.text.y = element_text(family = "Arial"),
    text = element_text(family = "Arial")
  )

p

ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/excitatory_dotplot.svg",
  plot = p,
  width = 10,
  height = 5,
)
```
```{r}
ggsave(
  "excitatory_mouse_umaps/excitatory_dotplot.svg",
  plot = p,
  width = 10,
  height = 5,
)
```


```{r}
library(dplyr)

# Collector counts
collector_counts <- table(URL$collector)

URL$collector_counts <- paste0(
  URL$collector,
  " (",
  collector_counts[URL$collector],
  ")"
)

# Sample counts
sample_counts <- table(URL$Sample)

URL$sample_counts <- paste0(
  URL$Sample,
  " (",
  sample_counts[URL$Sample],
  ")"
)

p <- DimPlot(URL, group.by  = "sample_counts")

ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/tables/mouse_exc_sample_counts.png",
  plot = p,
  width = 7,
  height = 5,
)
p

p <- DimPlot(URL, group.by  = "collector_counts")

ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/tables/mouse_exc_collector_counts.png",
  plot = p,
  width = 7,
  height = 5,
)
p
```





```{r}
p <- DimPlot(
  URL,
  group.by = "branch",
  cols = c(
    "transition" = "lightblue",  # swapped
    "early" = "maroon",
    "late" = "darkgreen"         # swapped
  )
)

ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/allURL_branch.png",
  plot = p,
  width = 7,
  height = 5,
)

```


```{r}
p <- DimPlot(
  URL,
  group.by = "TimePoint",
)

ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/allURL_timepoint.svg",
  plot = p,
  width = 7,
  height = 5,
)

```

```{r}
p <- FeaturePlot(URL, features = "Pax6", max.cutoff = 2)

ggsave("C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/allURL_Pax6.png",
  plot = p,
  width = 7,
  height = 5)

p <- FeaturePlot(URL, features = "Lhx9", max.cutoff = 2)

ggsave("C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/allURL_Lhx9.png",
  plot = p,
  width = 7,
  height = 5)

```

```{r}
p <- DimPlot(
  URL,
  group.by = "DCN_status",
  cols = c(
    "DCN" = "blue",
    "Other" = "grey",
    "RL" = "lightblue"
  )
)

ggsave("C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/allURL_dcn_status.png",
  plot = p,
  width = 7,
  height = 5)

```



```{r}
p <- FeaturePlot(URL, features = "Lmx1a", max.cutoff = 2)

ggsave("C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/allURL_Lmx1a.png",
  plot = p,
  width = 7,
  height = 5)

p <- FeaturePlot(URL, features = "Meis2", max.cutoff = 2)

ggsave("C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/allURL_Meis2.png",
  plot = p,
  width = 7,
  height = 5)

p <- FeaturePlot(URL, features = "Zfhx3", max.cutoff = 2)

ggsave("C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/allURL_Zfhx3.png",
  plot = p,
  width = 7,
  height = 5)

p <- FeaturePlot(URL, features = "Reln", max.cutoff = 2)

ggsave("C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/allURL_Reln.png",
  plot = p,
  width = 7,
  height = 5)

```

```{r}

```{r}
combined_alltp$med_2p_branch_coarse <- recode(
  combined_alltp$SCT_snn_res.2,
  
  "0"  = "Int/Lat Prog", 
  "1"  = "Med",
  "2"  = "Med", 
  "3"  = "IntA",
  "4"  = "RL", 
  "5"  = "IntA/Lat",
  "6"  = "Med", 
  "7"  = "IntA/Lat",
  "8"  = "RL", 
  "9"  = "Int/Lat Prog",
  "10" = "IntP", 
  "11" = "IntA",
  "12" = "Lat",
  "13" = "IntP", 
  "14" = "Med",
  "15" = "Med",
  "16" = "IntP", 
  "17" = "Med",
  "18" = "IntP",
  "19"  = "Med",
  "20" = "IntA/Lat", 
  "21" = "Med",
  "22" = "Lat",
  "23" = "IntP", 
  "24" = "IntP",
  "25" = "IntP",
  "26" = "Med"
)

branch_pal <- c(
  "Int/Lat Prog" = "yellow",
  "Med"          = "#08519c",
  "IntA/Lat"     = "#fca800",
  "IntA"     = "#f7a2f9",
  "Lat"     = "#cf3731",
  "RL"           = "lightblue",
  #"IntP/IntA"  
  "IntP"         = "#78b958",
  #"Med_gly"      = "#17becf",
  "Early Medial"        = "#824819",
  "Late Medial"         = "blue",
  "Midbrain-fated Cells"  = "black",
  "Granule Cells+UBCs" = "purple"
)

# Start with existing annotation
combined_alltp$med_2p_branch <- combined_alltp$med_2p_branch

# Replace with subareas where available
combined_alltp$med_2p_branch[!is.na(combined_alltp$subareas)] <-
  combined_alltp$subareas[!is.na(combined_alltp$subareas)]

# Collapse medial subtypes
combined_alltp$med_2p_branch[combined_alltp$subareas == "MedL"] <-
  "Early Medial"

combined_alltp$med_2p_branch[combined_alltp$subareas == "Med"] <-
  "Late Medial"

# Override MedDL cells
combined_alltp$med_2p_branch[combined_alltp$final.clusters2 == "MedDL"] <-
  "Late Medial"

# Start with the coarse annotation
combined_alltp$med_2p_branch_new <- as.character(combined_alltp$med_2p_branch_coarse)

# Rename Med -> Late Medial
combined_alltp$med_2p_branch_new[
  combined_alltp$med_2p_branch_coarse == "Med"
] <- "Late Medial"

# Replace with Early Medial where appropriate
combined_alltp$med_2p_branch_new[
  combined_alltp$med_2p_branch == "Early Medial"
] <- "Early Medial"

combined_alltp$med_2p_branch_new[combined_alltp$subareas == "MedL"] <-
  "Early Medial"


# Check the result
#table(combined_alltp$med_2p_branch_new)
#table(combined_alltp$med_2p_branch, combined_alltp$med_2p_branch_new)

p <- DimPlot(
  combined_alltp,
  group.by = "med_2p_branch_coarse",
  cols = branch_pal,
  label = FALSE,
  repel = TRUE
)

p




ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/CNcell_type_nomed_split.png",
  plot = p,
  width = 7,
  height = 5,
)
p

p <- DimPlot(
  combined_alltp,
  group.by = "TimePoint",
  #cols = branch_pal,
  label = FALSE,
  repel = TRUE
)



ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/CNtimepoint.svg",
  plot = p,
  width = 7,
  height = 5,
)
p
```
```{r}
DimPlot(
  combined_alltp,
  group.by = "med_2p_branch_new",
  cols = branch_pal,
  label = FALSE,
  repel = TRUE
)
```

```


