---
title: "01_Process snATAC-seq with ArchR"
output: html_notebook
---

```{r message=F}
library(ArchR);library(cowplot);library(Seurat)
set.seed(123)
addArchRThreads(threads = 20)
```

```{r}
combined_alltp = readRDS("C:/Users/kebsc/Downloads/combined_alltp2-2025.rds")
scrna_sub <- subset(combined_alltp, TimePoint != "adult")
```



```{r message=F}
library(ArchR);library(cowplot);library(Seurat)
set.seed(123)
addArchRThreads(threads = 20)
```


```{r}
library(ArchR)
set.seed(123)

# Set genome (IMPORTANT!)
addArchRGenome("mm10")  # Must be set before project creation

# Create Arrow Files (with updated arguments)
inputFiles <- c(
  "C:/Users/kebsc/Downloads/GSM5393630_E12_snATAC_fragments.tsv.gz",
  "C:/Users/kebsc/Downloads/GSM5393631_E13_snATAC_fragments.tsv.gz",
  "C:/Users/kebsc/Downloads/GSM5393632_E14_snATAC_fragments.tsv.gz"
)

ArrowFiles <- createArrowFiles(
  inputFiles = inputFiles,
  sampleNames = c("E12", "E13", "E14"),
  minTSS = 5,           # updated
  minFrags = 1250,      # updated
  maxFrags = 1e7,
  addTileMat = TRUE,
  addGeneScoreMat = TRUE,
  force = FALSE
)

```

# Infer doublets
For this dataset, we eventually decided not to remove doublets.


```{r}
proj <- ArchRProject(
  ArrowFiles = ArrowFiles,
  outputDirectory = "ArchR_E12_E14",
  copyArrows = TRUE
)

# Proceed with LSI, clustering, UMAP, etc.

```


# Perform dimensionality reduction and clustering
```{r}
proj <- addIterativeLSI(
  ArchRProj = proj,
  useMatrix = "TileMatrix", 
  name = "IterativeLSI", 
  iterations = 2, 
  clusterParams = list( #See Seurat::FindClusters
    resolution = c(0.2), 
    sampleCells = 10000, 
    n.start = 10), 
  varFeatures = 25000, 
  dimsToUse = 1:35,
  force = TRUE
)  

proj <- addClusters(
  input = proj,
  reducedDims = "IterativeLSI",
  method = "Seurat",
  name = "Clusters",
  resolution = 0.8,
  force = TRUE
)

# Visualizing in a 2D UMAP Embedding
proj <- addUMAP(
  ArchRProj = proj, 
  reducedDims = "IterativeLSI", 
  name = "UMAP", 
  nNeighbors = 30, 
  minDist = 0.3, 
  metric = "cosine",
  force = TRUE
)
```


```{r}
p1 <- plotEmbedding(ArchRProj = proj, colorBy = "cellColData", name = "Clusters", embedding = "UMAP")
p2 <- plotEmbedding(ArchRProj = proj, colorBy = "cellColData", name = "Sample", embedding = "UMAP")
#p2 <- plotEmbedding(ArchRProj = proj, colorBy = "cellColData", name = "label", embedding = "UMAP")

# plotPDF(p1,p2, name = "Plot-UMAP-Clusters_new.pdf", ArchRProj = proj, addDOC = FALSE, width = 10, height = 5)
```


```{r}
plotEmbedding(
  ArchRProj = proj,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Olig2",             # replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP")

plotEmbedding(
  ArchRProj = proj,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Zfhx3",             # replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP")

plotEmbedding(
  ArchRProj = proj,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Meis2",             # replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP")

plotEmbedding(
  ArchRProj = proj,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Ntng1",             # replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP")

plotEmbedding(
  ArchRProj = proj,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Lmx1a",             # replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP")

plotEmbedding(
  ArchRProj = proj,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Slc17a6",             # replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP")

plotEmbedding(
  ArchRProj = proj,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Gad1",             # replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP")

plotEmbedding(
  ArchRProj = proj,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Robo3",             # replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP")

plotEmbedding(
  ArchRProj = proj,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Atoh1",             # replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP")
```
```{r}
# Ensure the Clusters column exists
head(getCellColData(proj, "Clusters"))

# Get cluster cells (as character vector)
cluster_cells <- rownames(subset(getCellColData(proj), Clusters %in% c("C14", "C25", "C24")))

# Try subsetting again
proj_subset <- subsetArchRProject(
  ArchRProj = proj,
  cells = cluster_cells,
  outputDirectory = "Cluster_Subset_14_25_24_23",
  dropCells = TRUE,
  force = TRUE
)


```


```{r}
proj_subset <- addIterativeLSI(
  ArchRProj = proj_subset,
  useMatrix = "TileMatrix", 
  name = "IterativeLSI", 
  iterations = 2, 
  clusterParams = list( #See Seurat::FindClusters
    resolution = c(0.2), 
    sampleCells = 10000, 
    n.start = 10), 
  varFeatures = 10000, 
  dimsToUse = 1:20,
  force = TRUE
)  

proj_subset <- addClusters(
  input = proj_subset,
  reducedDims = "IterativeLSI",
  method = "Seurat",
  name = "Clusters",
  resolution = 0.4,
  force = TRUE
)

# Visualizing in a 2D UMAP Embedding
proj_subset <- addUMAP(
  ArchRProj = proj_subset, 
  reducedDims = "IterativeLSI", 
  name = "UMAP", 
  nNeighbors = 30, 
  minDist = 0.3, 
  metric = "cosine",
  force = TRUE
)
```


```{r}
plotEmbedding(ArchRProj = proj_subset, colorBy = "cellColData", name = "Clusters", embedding = "UMAP")
plotEmbedding(ArchRProj = proj_subset, colorBy = "cellColData", name = "Sample", embedding = "UMAP")

# plotPDF(p1,p2, name = "Plot-UMAP-Clusters_new.pdf", ArchRProj = proj, addDOC = FALSE, width = 10, height = 5)
```


```{r}
# Get cluster cells (as character vector)
cluster_cells <- rownames(subset(getCellColData(proj_subset), Clusters %in% c("C9", "C7", "C8", "C6", "C5")))

# Try subsetting again
proj_subset_cn <- subsetArchRProject(
  ArchRProj = proj_subset,
  cells = cluster_cells,
  outputDirectory = "Cluster_Subset_cn",
  dropCells = TRUE,
  force = TRUE
)

```
```{r}
proj_subset_cn <- addIterativeLSI(
  ArchRProj = proj_subset_cn,
  useMatrix = "TileMatrix", 
  name = "IterativeLSI", 
  iterations = 2, 
  clusterParams = list( #See Seurat::FindClusters
    resolution = c(0.2), 
    sampleCells = 10000, 
    n.start = 10), 
  varFeatures = 10000, 
  dimsToUse = 1:20,
  force = TRUE
)  

proj_subset_cn <- addClusters(
  input = proj_subset_cn,
  reducedDims = "IterativeLSI",
  method = "Seurat",
  name = "Clusters",
  resolution = 0.4,
  force = TRUE
)

# Visualizing in a 2D UMAP Embedding
proj_subset_cn <- addUMAP(
  ArchRProj = proj_subset_cn, 
  reducedDims = "IterativeLSI", 
  name = "UMAP", 
  nNeighbors = 30, 
  minDist = 0.3, 
  metric = "cosine",
  force = TRUE
)
```

```{r}
plotEmbedding(
  ArchRProj = proj_subset_cn,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Olig2",  
  size = 100, # replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP"
)
plotEmbedding(
  ArchRProj = proj_subset_cn,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Zfhx3", 
  size = 1,# replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP"
)
plotEmbedding(
  ArchRProj = proj_subset_cn,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Meis2", 
  size = 1,# replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP"
)
plotEmbedding(
  ArchRProj = proj_subset_cn,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Ntng1", 
  size = 1,# replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP"
)
plotEmbedding(
  ArchRProj = proj_subset_cn,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Lmx1a",             # replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP"
)
plotEmbedding(
  ArchRProj = proj_subset_cn,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Slc17a6", 
  size = 1,# replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP"
)
plotEmbedding(
  ArchRProj = proj_subset_cn,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Gad1",  
  size = 1,# replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP"
)
plotEmbedding(
  ArchRProj = proj_subset_cn,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Robo3",         
  size = 1,# replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP"
)
plotEmbedding(
  ArchRProj = proj_subset_cn,
  colorBy = "GeneScoreMatrix",    # or "GeneIntegrationMatrix"
  name = "Atoh1",  
  size = 1,# replace with your gene of interest, e.g., "CD3D"
  embedding = "UMAP"
)



```

```{r}
library(Seurat)
```


```{r}
scrna = readRDS("C:/Users/kebsc/Downloads/combined_alltp.rds")
```


```{r}
DefaultAssay(scrna) <- "SCT"
#scrna <- JoinLayers(scrna)
```

```{r}
scrna_sub = subset(scrna, subset = TimePoint %in% c("E12.5","E13.5", "E14.5"))
```

```{r}
scrna_sub <- FindNeighbors(scrna_sub, reduction = "mnn", dims = 1:30)
scrna_sub <- FindClusters(scrna_sub, resolution = 0.7)
```

```{r}
DefaultAssay(scrna_sub) <- "RNA"
scrna_sub <- JoinLayers(scrna_sub)
```


```{r}
DimPlot(scrna_sub, group.by = "seurat_clusters")
FeaturePlot(scrna_sub, features = "Lhx9")
```
```{r}
scrna_sub_intlat = subset(scrna_sub, seurat_clusters %in% c(4,6,9))

all_markers <- FindAllMarkers(
  object = scrna_sub_intlat,
  assay = "RNA",            # or "SCT", "integrated", etc.
  slot = "data",            # "data" for log-normalized data; "counts" for raw
  only.pos = TRUE,          # if TRUE, only return upregulated genes
  min.pct = 0.9,            # expressed in at least 10% of cells in either cluster
  logfc.threshold = 0.25    # minimum log2 fold-change
)

```

```{r}
library(dplyr)

top_markers <- all_markers %>%
  group_by(cluster) %>%
  top_n(n = 10, wt = avg_log2FC)   # top 10 markers per cluster

DoHeatmap(scrna_sub_intlat, features = unique(top_markers$gene)) + NoLegend()

```




```{r}
scrna_sub = FindVariableFeatures(scrna_sub, nfeatures = 300)
```
```{r}
proj_subset_cn <- saveArchRProject(ArchRProj = proj_subset_cn, outputDirectory = "Save-proj_subset_cn", load = TRUE)


```{r}
library(ArchR)
proj_subset_cn = loadArchRProject(path = "C:/Users/kebsc/Downloads/RNA_ATAC_Cerebellum-main/RNA_ATAC_Cerebellum-main/snATAC-seq/Save-proj_subset_cn")

```






```{r}
library(ArchR)
library(Seurat)

# 1. Extract gene names from ArchR GeneScoreMatrix
gsm <- getMatrixFromProject(proj_subset_cn, useMatrix = "GeneScoreMatrix")
genes_atac <- rowData(gsm)$name
genes_atac <- make.unique(genes_atac)

# 2. Extract gene names from Seurat RNA assay
genes_rna <- rownames(scrna_sub[["RNA"]])

# 3. Find shared genes
shared_genes <- intersect(genes_rna, genes_atac)
cat("Number of shared genes:", length(shared_genes), "\n")

# 4. Subset Seurat object to shared genes
scrna_sub <- subset(scrna_sub, features = shared_genes)

# Optional: Check after subsetting
cat("Number of genes in Seurat after subsetting:", nrow(scrna_sub[["RNA"]]), "\n")
```






```{r}
scrna_sub <-FindClusters(scrna_sub, resolution = 0.8)
DimPlot(scrna_sub, group.by = "seurat_clusters", label = TRUE)
```


```{r}
plotEmbedding(
  ArchRProj = proj_subset_cn,
  colorBy = "cellColData",
  name = "Sample",
  embedding = "UMAP",
  size = 2.5,
  returnGG = TRUE
)
```
```{r}
proj_subset_cn <- addClusters(
  input = proj_subset_cn,
  reducedDims = "IterativeLSI",
  name = "Clusters1",
  resolution = 0.6,
  force = TRUE# you can adjust (0.2–2.0)
)
```
```{r}
plotEmbedding(
  ArchRProj = proj_subset_cn,
  colorBy = "cellColData",
  name = "Clusters1",
  embedding = "UMAP",
  size = 2.5
)
```


```{r}
newNames <- c(
  "C1" = "Granule Cells",

  "C4" = "RL",
  "C5" = "Med",
  "C6" = "Med",
  "C7" = "IntA/Lat",
  "C8" = "IntA/Lat",
  "C9" = "IntP",
  "C10" = "IntA/Lat",
  "C11" = "Interposed/Lateral early"
  
  # add more if needed
)

proj_subset_cn$Clusters_renamed <- mapLabels(
  proj_subset_cn$Clusters1,
  newLabels = newNames
)

p <- plotEmbedding(
  ArchRProj = proj_subset_cn,
  colorBy = "cellColData",
  name = "Clusters_renamed",
  embedding = "UMAP",
  size = 2.5,
  returnGG = TRUE
)

# Increase legend size
p + theme(
  legend.key.size = unit(1.5, "cm"),     # size of legend keys
  legend.text = element_text(size = 18), # size of labels
  legend.title = element_text(size = 20) # title size
)


```
```{r}
# View cluster names
table(proj_subset_cn$Clusters_renamed)

# Remove C7
proj_filtered <- proj_subset_cn[!(proj_subset_cn$Clusters1 %in% c("C1", "C2", "C3")), ]

proj_filtered <- addIterativeLSI(
  ArchRProj = proj_filtered,
  useMatrix = "TileMatrix",
  name = "IterativeLSI",
  force = TRUE
)
proj_filtered <- addUMAP(
  ArchRProj = proj_filtered,
  reducedDims = "IterativeLSI",
  name = "UMAP",
  nNeighbors = 20,
  minDist = 0.2,
  force = TRUE
)
```


```{r}
proj_filtered <- addClusters(
  input = proj_filtered,
  reducedDims = "IterativeLSI",
  name = "Clusters",
  resolution = 1,
  force = TRUE
)
```


```{r}
plotEmbedding(
  ArchRProj = proj_filtered,
  colorBy = "cellColData",
  name = "Clusters1",
  embedding = "UMAP",
  size = 2.5
)

plotEmbedding(
  ArchRProj = proj_filtered,
  colorBy = "GeneScoreMatrix",
  name = "Lmx1a",
  embedding = "UMAP",
  size = 5
)

plotEmbedding(
  ArchRProj = proj_filtered,
  colorBy = "GeneScoreMatrix",
  name = "Olig2",
  embedding = "UMAP",
  size = 5
)

plotEmbedding(
  ArchRProj = proj_filtered,
  colorBy = "GeneScoreMatrix",
  name = "Irx3",
  embedding = "UMAP",
  size = 5
)


plotEmbedding(
  ArchRProj = proj_filtered,
  colorBy = "cellColData",
  name = "Sample",
  embedding = "UMAP",
  size = 2.5
)

plotEmbedding(
  ArchRProj = proj_filtered,
  colorBy = "cellColData",
  name = "Clusters_renamed",
  embedding = "UMAP",
  size = 2.5
)

```


```{r}
proj_filtered <- addPeakSet(
  ArchRProj = proj_filtered,
  peakSet = pk,
  #force = TRUE
)
```


```{r}
proj_filtered <- addPeakMatrix(proj_filtered)
```

```{r}
library(Seurat)
library(ArchR)
library(ggplot2)
library(RColorBrewer)

# 5. Run addGeneIntegrationMatrix
proj_filtered <- addGeneIntegrationMatrix(
    ArchRProj = proj_filtered,
    useMatrix = "GeneScoreMatrix",
    matrixName = "GeneIntegrationMatrix",
    reducedDims = "IterativeLSI",
    seRNA = scrna_sub,
    addToArrow = FALSE,
    nGenes = 300,
    dims = 1:20,
    reduction = "cca",
    groupRNA = "branch",
    nameCell = "predictedCell_Un",
    nameGroup = "predictedGroup_Un",
    nameScore = "predictedScore_Un"
)

cat("Integration complete!\n")


# --- Step 1: Prepare cluster labels ---
```


```{r}
# Adjust Seurat cluster labels by adding 1 (if needed)
cluster_chars <- as.character(scrna_sub$seurat_clusters)
cluster_nums <- as.numeric(cluster_chars)
cluster_nums_plus1 <- cluster_nums + 1
scrna_sub$seurat_clusters1 <- as.character(cluster_nums_plus1)

# Get unique cluster labels from both datasets as character vectors
rna_clusters <- sort(unique(scrna_sub$branch))
atac_clusters <- sort(unique(as.character(proj_filtered$predictedGroup_Un)))

# --- Step 2: Define a unified palette for all clusters ---

all_clusters <- union(rna_clusters, atac_clusters)
n_all <- length(all_clusters)
shared_palette_all <- rep(brewer.pal(8, "Set2"), length.out = n_all)
names(shared_palette_all) <- all_clusters

# Subset palette for each dataset's clusters
pal_rna <- shared_palette_all[rna_clusters]
pal_atac <- shared_palette_all[atac_clusters]
```


```{r}
# --- Step 3: Plot scRNA-seq UMAP with consistent palette ---

p_rna1 <- DimPlot(
  scrna_sub,
  group.by = "branch",
  #label = TRUE,
  cols = pal_rna
)  
  #ggtitle("scRNA-seq clusters") +
  #theme_minimal(base_size = 14) +
  #theme(plot.title = element_text(face = "bold"))

print(p_rna1)

ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/CN_Lmx1b.png",
  plot = p_rna1,
  width = 7,
  height = 5,
)

# --- Step 4: Plot ArchR UMAP with consistent palette ---
```


```{r}

pal_atac <- c(
  "Int/Lat Prog" = "yellow",
  "Med"          = "blue",
  "IntA/Lat"     = "#fca800",
  "IntA"     = "#f7a2f9",
  "Lat"     = "#cf3731",
  "RL"           = "lightblue",
  #"IntP/IntA"  
  "IntP"         = "#78b958",
  #"Med_gly"      = "#17becf",
  "Early Medial"        = "#824819",
  "Late Medial"         = "blue",
  "Midbrain-fated Cells"  = "black",
  "Granule Cells+UBCs" = "purple"
)

p_atac1 <- plotEmbedding(
  ArchRProj = proj_filtered,
  colorBy = "cellColData",
  name = "predictedGroup_Un",
  embedding = "UMAP",
  pal = pal_atac,
  returnGG = TRUE,
  size = 2.5
) 
  #ggtitle("scATAC-seq clusters") +
  #theme_minimal(base_size = 14) +
  #theme(plot.title = element_text(face = "bold"))


# Increase point size by modifying the geom_point layer
p_atac1 + 
  theme(
    legend.key.size = unit(1.4, "cm"),     # bigger legend boxes
    legend.text = element_text(size = 10), # bigger label text
    legend.title = element_text(size = 10) # bigger title
  ) +
  guides(color = guide_legend(override.aes = list(size = 3)))  # bigger legend points

print(p_atac1)

ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/scATAC_celltypes.png",
  plot = p_atac1,
  width = 7,
  height = 5,
)

```



```{r}
p_tss <- plotGroups(
  ArchRProj = proj_filtered,
  groupBy = "Sample",
  colorBy = "cellColData",
  name = "TSSEnrichment",
  plotAs = "violin",
  addBoxPlot = TRUE
) +
  geom_jitter(
    width = 0.15,
    size = 0.4,
    alpha = 0.35
  )
p_tss

ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/scATAC_TSS_enrichment_Sample.svg",
  plot = p_tss,
  width = 7,
  height = 5,
)

p_tss <- plotGroups(
  ArchRProj = proj_filtered,
  groupBy = "Sample",
  colorBy = "cellColData",
  name = "log10(nFrags)",
  plotAs = "violin",
  addBoxPlot = TRUE
) +
  geom_jitter(
    width = 0.15,
    size = 0.4,
    alpha = 0.35
  )
p_tss
ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/scATAC_nfrags_enrichment_Sample.svg",
  plot = p_tss,
  width = 7,
  height = 5,
)

p_tss <- plotGroups(
  ArchRProj = proj_filtered,
  groupBy = "predictedGroup_Un",
  colorBy = "cellColData",
  name = "TSSEnrichment",
  plotAs = "violin",
  addBoxPlot = TRUE
) +
  geom_jitter(
    width = 0.15,
    size = 0.4,
    alpha = 0.35
  )
p_tss

ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/scATAC_TSS_enrichment_clusters.svg",
  plot = p_tss,
  width = 7,
  height = 5,
)

p_tss <- plotGroups(
  ArchRProj = proj_filtered,
  groupBy = "predictedGroup_Un",
  colorBy = "cellColData",
  name = "log10(nFrags)",
  plotAs = "violin",
  addBoxPlot = TRUE
) +
  geom_jitter(
    width = 0.15,
    size = 0.4,
    alpha = 0.35
  )
p_tss

ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/scATAC_nfrags_enrichment_clusters.svg",
  plot = p_tss,
  width = 7,
  height = 5,
)

```



```{r}
p <- plotEmbedding(
  ArchRProj = proj_filtered,
  colorBy = "cellColData",
  name = "Sample",
  embedding = "UMAP",
  #pal = pal_atac,
  returnGG = TRUE,
  size = 3
)

# Increase point size by modifying the geom_point layer
p + 
  theme(
    legend.key.size = unit(1.4, "cm"),     # bigger legend boxes
    legend.text = element_text(size = 10), # bigger label text
    legend.title = element_text(size = 10) # bigger title
  ) +
  guides(color = guide_legend(override.aes = list(size = 3)))  # bigger legend points

ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/scATAC_timepoint.png",
  plot = p,
  width = 7,
  height = 5,
)

```


```{r}
p <- plotEmbedding(
  ArchRProj = proj_filtered,
  colorBy = "GeneScoreMatrix",
  name = "Lmx1a",
  embedding = "UMAP",
  returnGG = TRUE,
  size =2,
  plotAs = 'points',
  rastr = FALSE
)

# Increase point size by modifying the geom_point layer
p
ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/scATAC_Lmx1a.png",
  plot = p,
  width = 7,
  height = 5,
)
p <- plotEmbedding(
  ArchRProj = proj_filtered,
  colorBy = "GeneScoreMatrix",
  name = "Olig2",
  embedding = "UMAP",
  returnGG = TRUE,
  size =2,
  plotAs = 'points',
  rastr = FALSE)

# Increase point size by modifying the geom_point layer
p
ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/scATAC_Olig2.png",
  plot = p,
  width = 7,
  height = 5,
)
p <- plotEmbedding(
  ArchRProj = proj_filtered,
  colorBy = "GeneScoreMatrix",
  name = "Irx3",
  embedding = "UMAP",
  #pal = pal_atac,
  returnGG = TRUE,
  size =2,
  plotAs = 'points',
  rastr = FALSE
)

# Increase point size by modifying the geom_point layer
p#+ geom_point(size = 10)
ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/scATAC_Irx3.png",
  plot = p,
  width = 7,
  height = 5,
)
```
```{r}
# --- EDIT THESE ---
PEAK_BED  <- "C:/Users/kebsc/Downloads/mm10_bedfile.bed"
proj <- proj_filtered
suppressPackageStartupMessages({
  library(ArchR)
  library(rtracklayer)
  library(GenomicRanges)
  library(GenomeInfoDb)
  library(Matrix)
})



raw_lines <- readLines(PEAK_BED, warn=FALSE)
cat("Lines total:", length(raw_lines), "\n",
    "Lines starting with 'chr':", sum(grepl("^chr", raw_lines)), "\n")

# Import without lossy steps
pk0 <- tryCatch(import(PEAK_BED), error=function(e) NULL)
if (is.null(pk0)) {
  dt <- fread(text=paste(raw_lines[grepl("^chr", raw_lines)], collapse="\n"),
              header=FALSE, sep="\t", data.table=FALSE)
  names(dt)[1:3] <- c("chr","start","end")
  pk0 <- GRanges(dt$chr, IRanges(as.integer(gsub("[^0-9-]","",dt$start)),
                                  as.integer(gsub("[^0-9-]","",dt$end))))
}
cat("After import:", length(pk0), "\n")

pk1 <- pk0
seqlevelsStyle(pk1) <- "UCSC"
cat("After style:", length(pk1), "\n")

pk2 <- keepStandardChromosomes(pk1, pruning.mode="coarse")
pk2 <- pk2[seqnames(pk2)!="chrM"]
cat("After keepStandardChromosomes:", length(pk2), "\n")

# IMPORTANT: do NOT reduce yet
# If you previously resized+reduce, that can collapse thousands to a handful.

# pk2 from your last step (339,815 intervals)
pk <- pk2

# Give unique names if missing/duplicated
if (is.null(mcols(pk)$name)) mcols(pk)$name <- sprintf("peak_%06d", seq_along(pk))
mcols(pk)$name <- make.unique(as.character(mcols(pk)$name))
```

```{r}
# Register and build PeakMatrix
proj <- addPeakSet(proj, peakSet = pk, force = TRUE)
proj <- addPeakMatrix(proj)
```


```{r}
proj <- addMotifAnnotations(
  ArchRProj = proj,
  motifSet = 'cisbp',
  name = "Motif",
  #force = TRUE
)
```


```{r}
proj <- addBgdPeaks(proj, method = "ArchR")
```


```{r}
proj <- addDeviationsMatrix(
  ArchRProj = proj,
  peakAnnotation = "Motif",
  matrixName = "MotifMatrix",
  #force=TRUE
)

```

```{r}
motif_mat <- getMatrixFromProject(proj, useMatrix = "MotifMatrix")
dev_mat <- assays(motif_mat)[['z']]

smad_row <- grep("Rara", rownames(dev_mat), value = TRUE, ignore.case = TRUE)[1]

cells <- rownames(proj@cellColData)
proj$Rara_activity <- dev_mat[smad_row, cells]

```


```{r}
p <- plotEmbedding(
  ArchRProj = proj,
  colorBy = "cellColData",
  name = "Rara_activity",
  embedding = "UMAP",
  #pal = pal_atac,
  returnGG = TRUE,
  size =2,
  plotAs = 'points',
  rastr = FALSE
)
p
ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/scATAC_Rara_sites.png",
  plot = p,
  width = 7,
  height = 5,
)

plotEmbedding(
  ArchRProj = proj,
  colorBy = "GeneScoreMatrix",
  name = "Rara",
  embedding = "UMAP",
  #pal = pal_atac,
  returnGG = TRUE,
  size =2,
  plotAs = 'points',
  rastr = FALSE
)


```




```{r}
library(ggplot2)
df$predictedGroup_Un <- factor(
  df$predictedGroup_Un,
  levels = c(
    "Lat",
    "IntA",
    "IntA/Lat",
    "Int/Lat Prog",
    "IntP",
    "Med",
    "RL"
    
  )
)

library(ggplot2)

p_bar <- ggplot(
  df,
  aes(x = predictedGroup_Un, y = Rara_dev)
) +
  stat_summary(
    fun = mean,
    geom = "col",
    width = 0.45,
    fill = "grey80",
    color = "black",
    linewidth = 0.4
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.15,
    linewidth = 0.4
  ) +
  geom_jitter(
    width = 0.12,
    size = 0.5,
    alpha = 0.25,
    color = "black"
  ) +
  labs(
    x = NULL,
    y = "RARA motif activity"
  ) +
  theme_classic(base_size = 12) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1,
      colour = "black"
    ),
    axis.text.y = element_text(colour = "black"),
    axis.title.y = element_text(face = "bold"),
    axis.line = element_line(linewidth = 0.4),
    axis.ticks = element_line(linewidth = 0.4)
  )

p_bar


p_bar
ggsave(
  "C:/Users/kebsc/OneDrive - Johns Hopkins/Manuscripts/figs/excitatory_mouse_umaps/scATAC_Rara_diff_activity_clusters.svg",
  plot = p_bar,
  width = 7,
  height = 5,
)


```

```{r}
library(ggplot2)
library(ggpubr)
library(ArchR)


# -------------------------------
gene_mat <- getMatrixFromProject(proj_filtered, useMatrix = "GeneScoreMatrix")
gs_mat <- assays(gene_mat)[["GeneScoreMatrix"]]
```



```{r}
library(ArchR)
library(ggplot2)
library(ggpubr)
library(patchwork)

# -------------------------------
# 1. Extract Rara motif activity
# -------------------------------

# -------------------------------
# 2. Extract Rara GeneScore
# -------------------------------
gene_mat <- getMatrixFromProject(proj, useMatrix = "GeneScoreMatrix")
gene_names <- rowData(gene_mat)$name
cell_names_gene <- colData(gene_mat)$name

# Find Rara gene
gene_row <- grep("Rara", gene_names, ignore.case = TRUE)
if(length(gene_row) == 0) stop("Rara gene not found!")
```


```{r}
# Extract numeric vector and match to ArchR cells
rara_gene_vec <- assays(gene_mat)[["GeneScoreMatrix"]][gene_row, ]
#rara_gene_vec <- rara_gene_vec[match(proj_filtered$cellNames, cell_names_gene)]
proj_filtered$Rara_gene <- rara_gene_vec

# -------------------------------
# 3. Prepare data frame for plotting
# -------------------------------
df <- data.frame(
  predictedGroup_Un = as.factor(proj_filtered$predictedGroup_Un),
  Rara_gene = proj_filtered$Rara_gene
)

# All pairwise comparisons
comparisons_list <- combn(levels(df$predictedGroup_Un), 2, simplify = FALSE)

# -------------------------------
# 4. Create side-by-side plots
# -------------------------------

# GeneScore plot
ggplot(df, aes(x = predictedGroup_Un, y = Rara_gene, fill = predictedGroup_Un)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.15, size = 1, alpha = 0.6) +
  
  labs(y = "Rara GeneScore", x = "") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
groups = c("IntA/Lat", "IntP")
meta = getCellColData(proj)
cells_keep = rownames(meta)[meta$predictedGroup_Un %in% groups]

proj_sub = subsetArchRProject(ArchRProj = proj, cells = cells_keep,   force = TRUE)
```


```{r}
res <- getMarkerFeatures(ArchRProj = proj_sub, useMatrix = "MotifMatrix", groupBy = "predictedGroup_Un", testMethod = 'wilcoxon')
```


```{r}
library(SummarizedExperiment)
eps = 1e-6
M = assay(res, "Mean")
B = assay(res, "MeanBGD")
M <- as.matrix(assay(res, "Mean"))
B <- as.matrix(assay(res, "MeanBGD"))

L2F <- log2((M+eps)/(B+eps))
L2F[!is.finite(L2F)] = 0
L2F <- as.matrix(L2F)
dimnames(L2F)=dimnames(assay(res, "Mean"))
assays(res)[["Log2FC"]] <- L2F

k <- 10
mg <- plotMarkerHeatmap(seMarker = res,  cutOff = "FDR <= 0.05 & Log2FC >= 2.5", transpose = TRUE,plotLog2FC = TRUE, nLabel = 10)
mg
```
```{r}
#TF footprinting for RARE motif
group_col = "predictedGroup_Un"
# coverage per group (if not already present)
proj <- addGroupCoverages(
  ArchRProj = proj,
  groupBy   = group_col
)
```


```{r}
proj <- addMotifAnnotations(
  ArchRProj = proj,
  motifSet = "cisbp",
  annoName = "Motif",
  species = "Mus musculus",
  force = TRUE
)

motif_positions <- getPositions(
  ArchRProj = proj,
  name = "Motif"
)
```


```{r}
rar_motifs <- grep(
  "RAR|RARA|RARB|RARA|RARG|RXRA|RXRA|RXRB|RXRG",
  names(motif_positions),
  value= TRUE,
  ignore.case = TRUE)


rar_positions <- motif_positions[rar_motifs]

# compute footprints
seFoot <- getFootprints(
  ArchRProj = proj,
  groupBy   = group_col,
  positions = rar_positions
  #motifSet  = "Motif",
  #name      = rara_motif
)
```


```{r}
seFoot_IntP <- getFootprints(
  ArchRProj = proj,
  positions = rar_positions,
  groupBy = group_col,
  useGroups = c("IntP", "IntA", "Lat")
)
# plot RAR footprints across groups
p_fp <- plotFootprints(
  seFoot    = seFoot_IntP,
  ArchRProj = proj,
  #groupBy   = group_col,
  names = rar_motifs,
  smoothWindow = 20
)
p_fp

```

